---
title: "VOCA_data_and_programs"
author: "Kelsey"
date: "June 22, 2020"
output: html_document
bibliography: biblio.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(sp)
ut_crs <- "+proj=utm +zone=12 +datum=NAD83"
ut_shape <- st_read("C:/Users/Kelsey/Box/Utah VOC Project/R/Counties.shp")
ut_shape <- st_transform(ut_shape, ut_crs)
ut_data <- read.csv("C:/Users/Kelsey/Box/Utah VOC Project/Crime_by_county.csv")
ut_data <- merge(ut_shape, ut_data, by.x = "FIPS_STR", by.y = "GEOID")
colnames(ut_data)[which(colnames(ut_data) == "ï..County"):which(colnames(ut_data)=="Domestic.Violence")] <- 
  c("County_Name", "Population","Simple_Assault","Kidnapping","Sexual_Assault","Statutory_Rape",
    "Rape_Ind","Agg_Assault_Ind","Child_Abuse","Dom_Viol")

agency_data <- read.csv("C:/Users/Kelsey/OneDrive/Documents/UOVC Project/Organizations.csv")
agency_sf <- st_as_sf(agency_data, coords = c("Longitude", "Latitude"), crs = 4326)
agency_sf <- st_transform(agency_sf, crs=26912)
```

The purpose of this document is to show where the population centers of Utah are, where victims are most likely to be located, and where victims' services centers are.

## Sexual Assaults
@victcost gathers estimates of the societal costs of crime from 6 different studies and provides their own estimates. The cost per victimization of a sexual assault ranges from \$98,000 to \$450,000 (costs inflation-adjusted to 2020 using CPI). This includes tangible costs, like medical and mental health care spending, and intangible costs, like pain, suffering, and reduced quality of life. Each of the numbers were calculated using different methods, like the "jury compensation" approach or willingness to pay, and so varied widely.

From 2009 to 2018, there have been an average of 1,086 reported rapes each year (@utah_crime_2018). 

The following map shows the number of sexual assaults per county and the location of agencies that provide services to sexual assault victims. You can hover over a area or point to see the name of the county or agency, and additional information will show if you click on it.

```{r, echo = FALSE}
ut_data$SA_Rate <- 1000*ut_data$Sexual_Assault/ut_data$Population
agency_sf$awards_2019 <- agency_sf$SASP2019 + 
                          agency_sf$VAWA2019 + 
                          agency_sf$VOCA2019 + 
                          agency_sf$VOCAHousing2019
agency_sf$pct_SA <- agency_sf$AdultSexAssault/agency_sf$TotalIndviduals
agency_sf$pct_DV <- agency_sf$DomesticViolence/agency_sf$TotalIndviduals
agency_sf$DVSA <- 0
agency_sf$DVSA[agency_sf$OrgType %in% c("Domestic and Family Violence Organization", 
                                        "Sexual Assault Services organization (e.g., rape crisis center)",
                                        "Organization Provides Domestic and Family Violence and Sexual Assault Services", 
                                        "Campus-based victims services" )] <- 1
  
agency_sf$award_per_served <- agency_sf$awards_2019/agency_sf$TotalIndviduals

library(tmap)
tmap_mode("view")
tm_shape(ut_data) +
  tm_polygons("Sexual_Assault", 
              style = "kmeans", 
              id = "County_Name",
              popup.vars = c("Number of Sexual Assaults" = "Sexual_Assault", 
                             "Population" = "Population",
                             "Rate per 100,000" = "SA_Rate"),
              title = "Num. of Sex. Assaults") +
  tm_shape(agency_sf[agency_sf$DVSA == 1,]) + 
  tm_symbols(size = 0.1,
             col = "midnightblue",
             popup.vars = c("Total Individuals Served in 2019" = "TotalIndviduals",
                            "Total Awards in 2019"= "awards_2019",
                            "Percent Sexual Assault" = "pct_SA",
                            "Percent Domestic Violence" = "pct_DV",
                            "Funding Dollars Per Victim Served" = "award_per_served"
                            ))

```



## Domestic Violence
According to @victcost , the cost per victim of a sexual assault is roughly $157,000 (costs inflation-adjusted to 2020 using CPI).

## Child Abuse


## Where is funding going?
The following map shows where funding is going based on where agencies indicate they serve. Many agencies serve more than one county due to the rural nature of many counties in Utah, so funds are assumed to be distributed to counties based on the populations of where they serve. This may be unrealistic, as an agency likely serves the local community better than nearby counties.
```{r, echo=FALSE}
funding_data <- read.csv("C:/Users/Kelsey/Box/Utah VOC Project/Funding_by_county.csv")
funding_data <- merge(ut_shape, funding_data, by.x = "FIPS_STR", by.y = "GEOID")
colnames(funding_data)[colnames(funding_data)=="ï..County"] <- "County_name"

tm_shape(funding_data) +
  tm_polygons(c("Distributed_Awards", "POP_CURRES"), 
              style = "kmeans", 
              n = 8,
              id = "County",
              popup.vars = c("County" = "County_name",
                             "Award" = "Distributed_Awards",
                             "Award per Capita" = "Distributed_Awards_per_pop",
                             "Population" = "POP_CURRES"),
              title = c("Total Award", "Population")) +
  tm_facets(sync = TRUE, ncol = 2)

```
Funding is generally going where much of the population is. However, the amount of funding per capita is significantly higher in the most rural counties because there are fixed costs to running a shelter or agency that are spread across a smaller population. For example, the funding per capita in Salt Lake County is \$6, whereas in Iron County it is \$29, and in San Juan County its \$50.




## References
